---
title: "p8105_hw3_glj2117"
output: github_document
date: '2022-10-15'
---

```{r setup, include=FALSE}


library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1- Instacart data 

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


```{r}
library(p8105.datasets)
  data("instacart")

nrow(instacart) #1384617
ncol(instacart) #15

instacart %>% 
  group_by(aisle_id) %>% 
  count() %>% 
  print(n = 150) #134 aisles

```

## Problem 2
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

```
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```
```{r}
accel_data = read.csv("accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_post_midnight",
    names_prefix = "activity_",
    values_to = "activity_count") %>% 
 mutate(day = recode(day, "Monday" = 1, "Tuesday" = 2, "Wednesday" = 3, "Thursday" = 4, "Friday" = 5, "Saturday" = 6, "Sunday" = 7)) %>% 
  mutate(weekend_vs_weekday = if_else(day >= 6, "weekend", "weekday"))


```
The resulting dataset includes `r nrow(accel_data)` rows of data observations and `r ncol(accel_data)` variables. The data span 29 days over 5 weeks. With `day_id` indicating the day (1-29), day indicating the day of the week (where 1 = Monday and 7 = Sunday), `minute_post_midnight` indicating the time during the day at which the observation was taken, and `activity_count` being the metric of activity during that given minute of time. The `weekend_vs_weekday` column simply denotes whether the data were collected on a weekday or on a weekend day.

```
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_data %>% 
  group_by(day, day_id) %>% 
  summarize(sum(activity_count)) %>% 
 knitr::kable(digits = 1)

```
The trends days 4, 10, 16, 29 have the highest activity counts (in the 600,000s, while most of the other days fall in the ~200,000-400,000 range). There are two drastically low days for activity count, day 24 an day 31 (both within the last two weeks of the data collection). In terms of patterns by weekday, the weekdays seem to have realtively stable/similar activity counts, while increasing a bit on the weekends (except fot those two Saturdays with low points in week 5).
```
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_data %>% 
  ggplot(aes(x = minute_post_midnight, y = activity_count)) +
  geom_line(aes(color = day)) +
  viridis::scale_color_viridis(
    name = "day" 
 )

```
It is difficult to make many conclusions based on this graph, as it extremely busy, due to including so many datapoints. However, it is interesting to see (via the color distribution) thatactivity count stays relatively stable throughout the week and throughout the day. The outliers to the stability seem to be day 7 (Sunday) towards the end of the days, and a few mid-week points (days 2-4) in the beginning of the days. 

## Problem 3

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):
```{r}
library(p8105.datasets)
data(ny_noaa)

ny_noaa %>% 
  janitor::clean_names()
view(ny_noaa)


```
```
The dataset is `r nrow(ny_noaa)` observations long, and `r ncol(ny_noaa)` variables wide. Data are missing for all variables for early November 2007, as well as late July/early August 2008, and intermittently betwwen August and October of 2009. Additionally, tmin and tmax are missing for many dates, which can impact the accuracy of some of the summary statistics (such as mean) we want to calculate. 

```
Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}

ny_noaa = ny_noaa %>% 
  separate(col = 'date', into = c('year', 'month', 'day'), sep = '-')

ny_noaa %>% 
   group_by(snow) %>% 
    summarize(n_obs = n()) %>% 
    arrange(desc(n_obs))

```
The most commonly observed values for snowfall are 0 mm, 25 mm, and 13 mm. There are also 381221 NA values for snowfall, the second-most behind 0 mm. This is likely because there are many days without snowfall in the summer months, or where snowfall went unreported. 
```
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
ny_noaa = ny_noaa %>% 
  mutate(tmax = as.numeric(tmax))

ny_noaa = ny_noaa %>% 
  mutate(tmin = as.numeric(tmin))

month_tmax_min = ny_noaa %>% 
  group_by(id, year, month) %>% 
  summarize(
  mean_tmax = mean(tmax, NaN.rm = TRUE), 
  mean_tmin = mean(tmin, NaN.rm = TRUE)
)


month_tmax_min %>% 
  filter(month %in% c("01", "07")) %>% 
    ggplot(aes(x = year, y = mean_tmax, group = id)) + 
    geom_line(aes(color = id), alpha = .5) + 
    theme(legend.position = "none") + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    facet_grid(.~ month)
    

```
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
tmax_tmin =
  ny_noaa %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
  

tmax_tmin

snow_plot = ny_noaa %>% 
  filter(snow %in% (0:100)) %>% 
  ggplot(aes(x = snow, fill = year)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")


tmax_tmin + snow_plot
```


